---
title: "Dowjones time series with ARIMA"
author: "Rozenn Dahyot - https://www.scss.tcd.ie/Rozenn.Dahyot/"
date: "15 October 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Visualisation of  *dowjones* time series


```{r,warning=FALSE, message=FALSE}
require(fma)
tsdisplay(dowjones)
```

*dowjones* time series


- has a trend: the time plot shows an upward trend, the ACF shows very high correlations that are not decreasing quickly in amplitude as the lag $k$ increases; the PACF indicates   very high correlation between $y_t$ and $y_{t-1}$  but all other lags $k>1$ have PACF equivalent to 0.

- but no seasonality,

- and has a noise component.


Fitting a constant model to the time series:
$$
y_t=c+\epsilon_t
$$

and visualising the residuals $\lbrace \epsilon_t\rbrace_{1,cdots,n}$ as follow:
```{r,warning=FALSE, message=FALSE}
require(fma)
Arima(dowjones,order=c(0,0,0))
tsdisplay(Arima(dowjones,order=c(0,0,0))$residuals)
```


## Removing the trend in  *dowjones* by differencing 

To remove the trend, first order differencing is applied:

```{r,warning=FALSE, message=FALSE}
require(fma)
Arima(dowjones,order=c(0,1,0))
tsdisplay(Arima(dowjones,order=c(0,1,0))$residuals)
```

Note that the model fitted here is ($c=0$):
$$
y_t=y_{t-1}+\epsilon_t
$$

The PACF indicates PACF(1) to be non-zero whereas all other $PACF(k)=0$ for $k>1$. On the ACF a quick decrease to 0 value can be observed as $k$ increases.
A AR(1) model can be suggested to fit this residuals.

## Suggesting ARIMA(1,1,0)

```{r,warning=FALSE, message=FALSE}
require(fma)
Arima(dowjones,order=c(1,1,0))
tsdisplay(Arima(dowjones,order=c(1,1,0))$residuals)
```

Nothing more can be read with the time plot, ACF and PACF of the residuals.
So the final model identified using visualisation is 
$$
(y_t-y_{t-1})=\phi_1 (y_{t-1}-y_{t-2}) +\epsilon_t
$$

## Using AIC and BIC

Goodness of fit criteria associated with ARIMA models are given in the R output. 
We note that starting with ARIMA(0,0,0) model, to  ARIMA(0,1,0) model, and concluding with  ARIMA(1,1,0) model then  we have:

- a  decrease of BIC values (indicates an improvement in the goodness of fit of the model to the time series data)

- a decrease of AIC values (indicates an improvement in the goodness of fit of the model to the time series data) 

Using an automatic function in R a better model (based on AIC) is found ARIMA(1,1,1) (marginal improvement of AIC) but our ARIMA(1,1,0) is the best when BIC is chosen.   


```{r,warning=FALSE, message=FALSE}
require(fma)
auto.arima(dowjones,trace=TRUE)

Arima(dowjones,order=c(1,1,1))

```

**Exercises:** investigate R commands *Arima*, *arima*, *arima.sim*. Read https://otexts.com/fpp2/arima-r.html  


## Forecasts and prediction intervals

Once the best model is selected, it can be used for computing forecasts and prediction intervals:

```{r,warning=FALSE, message=FALSE}
require(fma)
plot(forecast(Arima(dowjones,order=c(1,1,0)),h=20,level = c(80, 95)))
forecast(Arima(dowjones,order=c(1,1,0)),h=20,level = c(80, 95))
```

