---
title: "Linear Regression"
author: "Rozenn Dahyot - https://Roznn.github.io"
date: "2020"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```








# Dataset


In this dataset (taken from: *An Introduction to Generalized Linear Models, A. J. Dobson & A. G. Barnett, 3rd edition p.96*), the response variable $y$ corresponds to the percentage of total calories obtained from complex carbohydrates for 20 male insulin-dependent diabetics who have been on a high carbohydrate diet for six months. Additional information is collected about the individuals taking part in the study including age (in years), weight (relative to ideal weight) and other calories intake from protein (as percentage).




```{r message=FALSE}

carbohydrate=c(33,40,37,27,30,43,34,48,30,38,50,51,30,36,41,42,46,24,35,37) # response vector
age=c(33,47,49,35,46,52,62,23,32,42,31,61,63,40,50,64,56,61,48,28)
weight=c(100,92,135,144,140,101,95,101,98,105,108,85,130,127,109,107,117,100,118,102)
protein=c(14,15,18,12,15,15,14,17,15,14,17,19,19,20,15,16,18,13,18,14)



```

# Fitting a linear model with R functions

The   R functions *lm()* and *glm()* can be used to fit a linear model of the form
$$
y_{\text{carbohydrate}}=\underbrace{\beta_0+\beta_1 \ x_{\text{age}}+\beta_2 \ x_{\text{weight}} + \beta_3 \ x_{\text{protein}}}_{\hat{y}}+\epsilon
$$
Once estimates for parameters $\beta$s are available, then a predictor $\hat{y}$ can be computed for any inputs $(x_{\text{age}},x_{\text{weight}},x_{\text{protein}})$.



```{r message=FALSE}
#using lm

res.lm=lm(carbohydrate~age+weight+protein)
summary(res.lm)


#using glm
res.glm=glm(carbohydrate~age+weight+protein,family=gaussian)
summary(res.glm)


```


# Fitting a linear model using linear algebra

This code used linear algebra to recover some of the results found by the R functions *lm()* and *glm()*. 



We define matrix $\mathrm{X}$ with the explanatory variables, and vector $\mathbf{y}$ collecting the responses.

```{r message=FALSE}
X=matrix(1,20,4)
X[,1]=c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)
X[,2]=age
X[,3]=weight
X[,4]=protein
Y=carbohydrate
```


## Least squares solution

The least square solution is computed with:
$$
\hat{\beta}=(\mathrm{X}^{T}\mathrm{X})^{-1} \ \mathrm{X}^{T} \mathbf{y}
$$

Using R, notice how we recover the same estimates computed with R function functions *lm()* and *glm()*:
```{r message=FALSE}

BetaHat=solve(t(X)%*%X)%*%t(X)%*%carbohydrate
BetaHat
```


## Residual standard error


The Residual Standard Error (RSE) reported by the funtion *lm()* can be also computed using linear algebra as follow:
$$
RSE=\sqrt{\frac{SSE}{n-p}}
$$
with $n$ is the number of observations and $p$ is the degree of freedom of the model (i.e. number of parameters $\lbrace \beta_0,\beta_1,\beta_2,\beta_3\rbrace$ to estimate in the model).
The Sum of Square Error (SSE) is computed as :
$$
SSE=\sum_{i=1}^n (y_{i}-\hat{y}_i)^2
$$


```{r message=FALSE}
fitted =X %*% BetaHat
residuals=carbohydrate-fitted
SSE=t(residuals)%*%residuals # Sum of Square Error
sigmahatsq=SSE/(length(carbohydrate)-length(BetaHat))
#Residual standard error as reported by lm()
sqrt(sigmahatsq)
```


## Standard Errors of estimated parameters


The uncertainty associated with the parameters $\beta$s (Std. Error as reported with *lm()* and *glm()*) can also be computed: 


```{r message=FALSE}
# uncertainty of BetaHat (standard error)
CovBetaHat=solve(t(X)%*%X)*as.numeric(sigmahatsq)
Std.Error.BetaHat=sqrt(diag(CovBetaHat))
Std.Error.BetaHat
```


## Multiple R-squared

Similarly R-squared (as reported with *lm()*) can be computed:
$$
R^2=\frac{\sum_{i=1}^n (\hat{y}_{i}-\bar{y})^2}{\sum_{i=1}^n (y_{i}-\bar{y})^2}
$$
using the sample mean $\bar{y}=\frac{1}{n}\sum_{i=1}^n y_{i}$.

```{r message=FALSE}
#Multiple R-squared: (coefficient of determination) 
S0=sum((carbohydrate-mean(carbohydrate))*(carbohydrate-mean(carbohydrate)))
Rsq=(S0-SSE)/S0
Rsq


S1=sum((fitted-mean(carbohydrate))*(fitted-mean(carbohydrate)))
Rsq2<-S1/S0
Rsq2

```

