---
title: "Time series *beer* with Holt-Winters algorithms"
author: "Rozenn Dahyot - https://www.scss.tcd.ie/Rozenn.Dahyot/ "
date: " 2019 "
output: 
   html_document: default
   
---

```{r setup, include=FALSE,warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



The time series *beer* is in the fma package:
```{r message=FALSE}
# make sure to type in Rstudio console: install.packages("fma")
require("fma")
beer

```
Description of beer: 
Monthly Australian beer production: Jan 1991 to Aug 1995.


## Visualisation of time series


### Time plot

You can also embed plots, for example the time plot:

```{r message=FALSE }
plot(beer)
```


### Time plot with ACF and PACF

The ACF and PACF can be displayed with the time plot:


```{r message=FALSE}
tsdisplay(beer,lag.max=70)
```

Other packages exist for visualisation.  For instance you may prefer 
using packages *ggplot2* 


```{r message=FALSE}
# make sure to type in Rstudio console: install.packages("ggplot2")
require(ggplot2)
ggtsdisplay(beer,lag.max = 70)
```

The timeplot shows seasonality, no real trend and some noise.  
For the time series *beer*, the ACF displays maxima at lags 12, 24, 36 (...) characteristic of seasonality with a period of 12 months (as it is  monthly data).



**Exercise:** Show how to use the R functions *acf()* and *pacf()* for the time series *beer*. Explain the the differences with *tsdisplay()* (hint: look at the x-axis)

**Exercise:** Comment on the information provided by the PACF for the time series *beer*.


### seasonplot()

The season plot confirms there is no trend.
```{r message=FALSE}
seasonplot(beer)
```

### Function *ts()*

Note that *beer* time series is stored in *fma* package explicitly with this information of periodicity 12. 
For the *mink* time series which has no frequency information stored, then you will need to use the function *ts* to indicate a period (if you see one!) for this time series e.g.:

```{r message=FALSE}
tsdisplay(ts(mink,freq=10))
seasonplot(ts(mink,freq=10))
```


### Function *decompose()*

Note that the function *decompose* can be visually misleading:

```{r message=FALSE}
plot(decompose(beer))
```

It may look like you have trend!

**Exercise: **Read chapter 6 at  
https://otexts.com/fpp2/decomposition.html .
What other R function can be used for decomposing time series  ?

## Fitting Holt Winters algorithms


Only seasonal algorithms are appropriate for fitting the time series *beer*.

```{r message=FALSE}
HoltWinters(beer, seasonal="additive")$SSE
HoltWinters(beer, seasonal="multiplicative")$SSE
```

Using SSE (Sum of Square Errors), the seasonal Holt Winters algorithm (SHWx) algorithm fits best the time series *beer* (lowest SSE). 

**Exercise:** How to compute MAPE with R ? 
Would the same Holt-Winters algorithm  selected with MAPE ?

## Computing forecasts 


```{r message=FALSE}
HoltWinters(beer, seasonal="multiplicative")
predict(HoltWinters(beer, seasonal="multiplicative"))
```

Using the formula
$$
F_{n+k} = (L_n+k \times b_n)\times S_{n+k-s}
$$
with $k=1$, $s=12$ (frequency for beer data), and with the notation used in the lecturenotes https://www.scss.tcd.ie/Rozenn.Dahyot/ST3010/RzDTimeSeriesForecasting.pdf
(chapter 8), the prediction can be computed (instead of using the R command *predict()*) as follow

```{r message=FALSE}
k=1  # prediction computed 1 month ahead
s=12 # frequency 12 month 
Ln=HoltWinters(beer, seasonal="multiplicative")$coefficients[1]
bn=HoltWinters(beer, seasonal="multiplicative")$coefficients[2]
snk= HoltWinters(beer, seasonal="multiplicative")$coefficients[3]
(Ln+k*bn)*snk
```

Likewise prediction can be computed for $k=5$ months ahead from the last observation of the time series *beer* : 

```{r message=FALSE}
k=5  # prediction computed 1 month ahead
s=12 # frequency 12 month 
Ln=HoltWinters(beer, seasonal="multiplicative")$coefficients[1]
bn=HoltWinters(beer, seasonal="multiplicative")$coefficients[2]
snk= HoltWinters(beer, seasonal="multiplicative")$coefficients[7]
(Ln+k*bn)*snk
predict(HoltWinters(beer, seasonal="multiplicative"),n.ahead=5)[5]
```


### Using *forecast()*

```{r message=FALSE}
forecast(HoltWinters(beer, seasonal="multiplicative"),h=1)
```

```{r message=FALSE}
plot(forecast(HoltWinters(beer, seasonal="multiplicative")))
```

**Exercise:** Explain how the R function  *forecast()* relates to *predict()*. Explain how  prediction intervals are computed. 



## Other questions

**Exercise:** What Warnings should one be aware of about using  the computed forecasts with Holt Winters algorithms ? What hypotheses are made when computing  the prediction intervals associated with the forecasts ?


**Exercise:** Read *The Holt-Winters Approach to Exponential Smoothing: 50 Years Old and Going Strong* by Paul Goodwin in FORESIGHT Fall 2010 (  http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.401.2999&rep=rep1&type=pdf
). Explain how Holt-Winters algorithms have been extended ?

